A discussion of the intersection of the kernels of each $(1+T_i)$ acting on $W_{2n+r}^r$

### 2019-07-28 tex/kernel.tex
First draft section 2, overview, bibliography, some more data
almost finished

In this section we will fully characterize $\cap\text{ker}(1+T_i)$ when $e=n+r+1$. Note that we still have not proved the kernel is trivial when $e$ divides but is not equal to $n+r+1$. That proof requires results from this section, and will come next section.

\vspace{5mm}
We will characterize our kernel in a recursive fashion. To motivate this characterization, consider the following lemma. In simpler terms, this lemma states that if a basis element $\beta$ has some sub-matching $\beta(a,b)$, we may act on $\beta$ as if we are acting on $\beta(a,b)$.

@@

	Generally, for any matching $\alpha$, define $x_\alpha$ to be the rational function corresponding to the coordinate of that basis element in its respective kernel element. We will need to prove this is well defined.

@@

	\item Use lemma 3.1 to finish the proof through induction.
\end{enumerate}

@@

>>>>

In this section we will fully characterize $\cap\text{ker}(1+T_i)$ when $e=n+r+1$. We will prove that it is one dimensional and give a basis. Note that we still have not proved the kernel is trivial when $e$ divides but is not equal to $n+r+1$. That proof requires results from this section, and will come next section.

@@

	Generally, for any matching $\alpha$, define $x_\alpha$ to be the rational function corresponding to the necessary coordinate of that basis element in its respective kernel element. These functions are defined recursively in the next statement.

@@

	\item Finish the proof through induction.
\end{enumerate}

@@

\vspace{5mm}
The following few corollaries will help to simplify some later proofs.

\begin{corollary}
	Let $w\in\cap\text{ker}(1+T_i)$, $w\not=0$. Suppose $\beta(1,a)$ is a sub-matching with no anchors. Then:

	$$x_\beta=x_{\beta(1,a)}(q)x_{\beta(a+1,2n+r)}(q)\frac{Q_{n+r-1}...Q_{n+r-a/2}}{Q_1...Q_{a/2-1}}$$
\end{corollary}

\begin{proof}
	Define $a_1=\beta(1)$, $a_i=\beta(a_{i-1}+1)$. Then for some $j$ we have $a_j=a$. If $j=1$, the statement is the same as the proposition. Suppose that the statement is true for any matching with $a_v=a$, $v<j$. Then the statement holds for the sub-matching $\beta(a_1+1,2n+r)$, and we have:

	$$x_\beta(q)=x_{\beta(1,a_1)}(q)x_{\beta(a_1+1,2n+r)}(q)\frac{Q_{n+r-1}...Q_{n+r-a_1/2}}{Q_1...Q_{a_1/2-1}}=$$

	$$x_{\beta(1,a_1)}(q)x_{\beta(a_1+1,a)}(q)x_{\beta(a+1,2n+r)}(q)\frac{Q_{n+r-1}...Q_{n+r-a_1/2}}{Q_1...Q_{a_1/2-1}}\frac{Q_{n+r-a_1/2-1}...Q_{n+r-a/2}}{Q_1...Q_{a/2-a_1/2-1}}=$$

	$$x_{\beta(1,a_1)}(q)x_{\beta(a_1+1,a)}(q)x_{\beta(a+1,2n+r)}(q)\frac{Q_{n+r-1}...Q_{n+r-a_1/2}}{Q_1...Q_{a_1/2-1}}\frac{Q_{n+r-a_1/2-1}...Q_{n+r-a/2}}{Q_1...Q_{a/2-a_1/2-1}}(\frac{Q_{a/2-1}...Q_{a/2-a_1/2}}{Q_{a/2-1}...Q_{a/2-a_1/2}})=$$

	$$x_{\beta(1,a_1)}(q)x_{\beta(a_1+1,a)}(q)x_{\beta(a+1,2n+r)}(q)\frac{Q_{a/2-1}...Q_{a/2-a_1/2}}{Q_1...Q_{a_1/2-1}}\frac{Q_{n+r-1}...Q_{n+r-a/2}}{Q_1...Q_{a/2-1}}=$$

	$$=x_{\beta(1,a)}(q)x_{\beta(a+1,2n+r)}(q)\frac{Q_{n+r-1}...Q_{n+r-a/2}}{Q_1...Q_{a/2-1}}$$

	This proves the inductive step, so with base case $j=1$ the corollary holds.

\vspace{5mm}
\begin{corollary}
	$\beta\in W_{2n+r}^r$, then $x_\beta(q)\not=0$ if $e>n+r$.
\end{corollary}

\begin{proof}
	For our base cases, if $2n+r=2$ all coefficients are 1, which is nonzero for any $q$.

	Assume the statement is true for all $2n'+r'<2n+r$. We have $$x_\beta(q)=x_{\alpha_1}(q)x_{\alpha_2}(q)\frac{Q_{n+r-1}...Q_{n+r-(a/2)}}{Q_1...Q_{a/2-1}}$$
	or $$x_\beta(q)=x_{\beta(2,2n+r)}(q)$$

	If $e>n+r$, non of the $Q_i$ terms are zero, and $n'+r'<n+r<e$ for any of the sub-matchings that appear, so those coordinates are nonzero and the corollary holds.

\vspace{5mm}
The previous proposition fully characterizes any possible kernel element when $Q_1...Q_{n+r-1}\not=0$. In particular, the following corollary holds:

\begin{corollary}
	When $Q_1..Q_{n+r-1}\not=0$ and the kernel is nontrivial, the kernel is one dimensional.
\end{corollary}

This corollary follows from the fact that we may write the coordinate of any basis element as proportional to the coordinate of the rainbow basis element.

\vspace{5mm}
\noindent\rule{16.5cm}{0.4pt}

\vspace{5mm}
The remainder of this section will be used to prove that when $e=n+r+1$, the element specified by proposition 3.2 is indeed an element of the kernel.

In service of this goal, a few lemmas will be helpful. The first lemma states that sub-matchings behave identically to the corresponding representations.

\vspace{5mm}
To verify the kernel element, we will need to know exactly which basis elements are mapped to a specific basis element by a given $(1+T_i)$. The next two lemmas help address this question.

\begin{lemma}
	Take some basis element $\beta\in W_{2n+r}^r$. Suppose $\beta(a)=b$ for some $b>a+1$, and that $(1+T_i)\beta=(1+q)\beta$ for some $a<i<b-1$. We then have a subrepresentation $\beta(a,b)$ and can define $Y_\beta$ as in the previous lemma. Then for all basis elements $\alpha$ such that $(1+T_i)\alpha=q^{1/2}\beta$, we have that $$\alpha\in Y_\beta$$.

	Similarly, if $\beta$ has some anchor at position $u$, and $(1+T_i)\beta=(1+q)\beta$ for some $i>u$, we again have a subrepresentation $\beta(u,2n+r)$ and may define $Y_\beta$ as before. Then for all basis elements $\alpha$ such that $(1+T_i)\alpha=q^{1/2}\beta$, we have that $\alpha\in Y_\beta$ again.
\end{lemma}

\begin{proof}
	This lemma follows from an observation I made in section 2: a transposition can only create two arcs or an arc and an anchor.

	For the first case of this lemma, if $\alpha\not\in Y_\beta$ either $\alpha(1,a-1)\not=\beta(1,a-1)$ or $\alpha(b+1,2n+r)\not=\beta(b+1,2n+r)$. Suppose it is the first case. Then for some $s,t\in [1,a-1]$, $s<t$, we have $\beta(s)=t$ and $\alpha(s)\not=t$. To have $(1+T_i)\alpha=q^{1/2}\beta$ we must have $\alpha(t)=i+1$, $\alpha(s)=i$. But then $\alpha(a)\not=b$ and $\alpha(a)\not=i$ or $i+1$, so $((1+T_i)\alpha)(a)\not=b$ and $(1+T_i)\alpha\not=q^{1/2}\beta$. The same argument proves the $\alpha(b+1,2n+r)\not=\beta(b+1,2n+r)$ case.

	An analogous argument proves the anchor case. Specifically, the anchor cannot exist at position $u$ and is not created by action of $(1+T_i)$ if $\alpha(s)=i$ and $\alpha(t)=i+1$.

\vspace{5mm}
It is important to note that lemma 3.6 only references cases where a transposition acts under an arc or to the right of an anchor. An example is given in figure 7.

The next lemma characterizes cases where the transposition is not under any arcs and all anchors are to the right.

\begin{figure}[b]
	\[
	\begin{tabular}{l l l l}
	\GeneralizedAction{13}{1/4,2/3,6/7,8/13,9/10,11/12}{1}{5/1}{5/4}{9}{1/4,2/3,6/7,8/13,9/10,11/12}{5/1}{(1+q)}\\

	\GeneralizedAction{13}{1/4,2/3,6/7,8/9,10/13,11/12}{1}{5/1}{5/4}{9}{1/4,2/3,6/7,8/13,9/10,11/12}{5/1}{(q^{1/2})}\\

	\GeneralizedAction{13}{1/4,2/3,6/7,8/13,9/10,11/12}{1}{5/1}{5/4}{6}{1/4,2/3,6/7,8/13,9/10,11/12}{5/1}{(1+q)}\\

	\GeneralizedAction{13}{1/4,2/3,5/6,8/13,9/10,11/12}{1}{7/1}{5/4}{6}{1/4,2/3,6/7,8/13,9/10,11/12}{5/1}{(q^{1/2})}\\

	\end{tabular}
	\]

	\caption{In the first line we act under an arc, so if another element without that arc is sent to that element, it must fix the arc as shown in the second line. In the third line we act to the right of an anchor, so if another element without that anchor is sent to that element, it must fix the anchor as shown in the fourth line. 
	}
	\label{Action}
\end{figure}

Essentially, this lemma states that the only elements sent to the same element are those which break at most one of the top level arcs to the left of the leftmost anchor, or that break the leftmost anchor. An illustration is given in figure 8.

\vspace{5mm}
\begin{lemma}
	Take a basis element $\beta\in W_{2n+r}^r$. Suppose the leftmost anchor in $\beta$ is at index $b$. Define $a_j$ such that $\beta(a_j)=a_{j-1}+1$ and $\beta(a_1)=1$ for all $j$ such that $a_j<b$.

	Suppose $(1+T_i)\beta=(1+q)\beta$ for some $i<b-1$ where $\not\exists s,t$ such that $\beta(s)=t$ and $s<i,t>i+1$. Suppose there is some basis element $\alpha$ such that $(1+T_i)\alpha=q^{1/2}\beta$. Then I claim the following:

	\vspace{5mm}
	\begin{center}
		(i) $\beta(a_{j-1}+2,a_j-1)=\alpha(a_{j-1}+2,a_j-1)$ for all $j$.

		\vspace{5mm}
		(ii) $\beta(b+1,2n+r)=\alpha(b+1,2n+r)$

		\vspace{5mm}
		(iii)
		If $b$ is not an anchor in $\alpha$, $\beta(a_j)=\alpha(a_j)$ for all $j$ such that $a_j\not=i+1$.

		\vspace{5mm}
		(iiii) If $b$ is an anchor in $\alpha$, there exists exactly one value of $j$ such that $\alpha(a_j)\not=\beta(a_j)$ and $a_j\not=i+1$
	\end{center}
\end{lemma}

\begin{proof}

	(i)

	Suppose that, for some $j$ there exists $s,t\in [a_{j-1}+2,a_j-1]$ such that $\beta(s)=t$ but $\alpha(s)\not=t$. Then if $(1+T_i)\alpha=q^{1/2}\beta$ we must have $\alpha(i)=s$ or $t$ and $\alpha(i+1)=s$ or $t$. But, by definition, $i,i+1\not\in[a_{j-1}+1,a_j]$, so this implies $\alpha(a_j)\not=a_{j-1}+1,i,i+1$, so $((1+T_i)\alpha)(a_j)\not=a_{j-1}+1$ and $(1+T_i)\alpha\not=q^{1/2}\beta$. So (i) is proved.

	\vspace{5mm}
	(ii)

	The proof of (ii) is analogous to the proof of (i). We cannot have $\beta(b+1,2n+r)\not=\alpha(b+1,2n+r)$ and $\beta(b+1,2n+r)=((1+T_i)\alpha)(b+1,2n+r)$ if $((1+T_i)\alpha)(b)=b$.

	\vspace{5mm}
	(iii)

	If $b$ is not an anchor in $\alpha$ and $(1+T_i)\alpha=q^{1/2}\beta$, we must have $i$ an anchor in $\alpha$, and $\alpha(i+1)=b$. No other nodes in $\alpha$ are changed, so this proves (iii).

	\vspace{5mm}
	(iiii)
	From (i)-(iii) we have that the only remaining matchings that can differ are the $(a_{j-1}+1,a_j)$ matchings. If one of them differs, by the same argument as before it must be fixed by the action of $(1+T_i)$, and no other nodes are changed, so (iiii) is proved.

\begin{figure}[b]
	\[
	\begin{tabular}{l l l l}
	\GeneralizedAction{15}{1/6,2/3,4/5,7/8,9/10,12/15,13/14}{1}{11/1}{5/4}{7}{1/6,2/3,4/5,7/8,9/10,12/15,13/14}{11/1}{(1+q)}\\

	\GeneralizedAction{15}{1/8,2/3,4/5,6/7,9/10,12/15,13/14}{1}{11/1}{5/4}{7}{1/6,2/3,4/5,7/8,9/10,12/15,13/14}{11/1}{(q^{1/2})}\\

	\GeneralizedAction{15}{1/6,2/3,4/5,7/10,8/9,12/15,13/14}{1}{11/1}{5/4}{7}{1/6,2/3,4/5,7/8,9/10,12/15,13/14}{11/1}{(q^{1/2})}\\

	\GeneralizedAction{15}{1/6,2/3,4/5,8/11,9/10,12/15,13/14}{1}{7/1}{5/4}{7}{1/6,2/3,4/5,7/8,9/10,12/15,13/14}{11/1}{(q^{1/2})}\\

	\end{tabular}
	\]

	\caption{The action of $(1+T_7)$ fixes the first basis element. Shown are all the basis vectors sent to the same element by the same transposition. Note that in all of them nodes 2-5 and 12-15 are the same. This illustrates (i) and (ii) in lemma 3.5. Note that in the last case where the anchor is in a different place, 1,6 and 9,10 are still matched. This illustrates (iii). In the middle two cases where the anchor is in the same place, only one of 1,6 or 9,10 are not paired. This illustrates (iiii). 
	}
	\label{Action}
\end{figure}

\vspace{5mm}
We are now ready to prove existence of a kernel element. To prove this, we will show that if $w\in W_{2n+r}^r$ is as characterized above, the coordinate of any basis element in $(1+T_i)w$ is zero. This will split into various cases related to the previous lemmas.

\vspace{5mm}
\begin{proposition}
	Suppose $e=n+r+1$. Then $\cap\text{ker}(1+T_i)\not=\emptyset$.
\end{proposition}

\begin{proof}

	Assume inductively that the statement holds for all $W_{2n'+r'}^{r'}$ where $2n'+r'<2n+r$.

	As a base case, when $2n'+r'\leq 2$, the representation is at most one dimensional. If the one basis element has only anchors, it is sent to zero by any $(1+T_i)$, and is in the kernel. If the single basis element is a single arc, it is sent to $(1+q)$ times itself, and we take $e=n+r+1=2$ so $1+q=0$ and the base case holds.

	\vspace{5mm}
	Now we will prove the inductive step. Take $w$ as recursively defined by proposition 3.1. Formally, suppose $x_\alpha$ is the rational function corresponding to the coordinate of $\alpha$ in its respective kernel element for $\alpha\in W_{2n'+r'}^{r'}$, $2n'+r'<2n+r$. Then we define $w$ by its coordinate vector: $\beta\in W_{2n+r}^r$, if $\beta(1)=a\not=1$ then the coordinate of $\beta$ is $x_{\beta(2,a-1)}(q)x_{a+1,2n+r}(q)\frac{Q_{n-1}...Q_{a/2}}{Q_1...Q_{a/2-1}}$; if $\beta(1)=1$ then the coordinate of $\beta$ is $x_{\beta(2,2n+r)}(q)$. 
	\\

	Let $E_\beta\subset W_{2n+r}^r$ be the pre-image of $\beta\in (1+T_i)W_{2n+r}^r$ under the action of $(1+T_i)$. To prove $w$ is in the kernel, we must show the following: 
	\begin{center}
	(i) $(1+q)x_\beta(q)+\sum_{\alpha\in E_\beta,\alpha\not=\beta}q^{1/2}x_\alpha(q)=0\text{ for all basis elements }\beta$
	\end{center}
	Inductively, we assume this equation holds for basis elements in smaller representations $W_{2n'+r'}^{r'}$, but only for $q$ such that $e=n'+r'+1$. For the following proof we will need a slightly stronger inductive assumption. Take $\beta'\in W_{2n'+r'}^{r'}$, and suppose either that $\beta'(1)=2n'+r'$, and that $T_i\beta'=(1+q)\beta'$, $1<i<2n'+r'-1$, or that 1 is an anchor in $\beta$. Defining $E_{\beta'}$ as before, we assume

	\begin{center}
		(ii) $(1+q)x_{\beta'}+\sum_{\alpha\in E_{\beta'},\alpha\not=\beta'}q^{1/2}x_\alpha=0\text{ for any } q \text{ with }e>n'+r'$
	\end{center}

	\vspace{5mm}
	To prove the inductive step for both (i) and (ii), we must split into cases:

	\begin{enumerate}
	\item Suppose $\beta\in (1+T_i)W_{2n+r}^r$ for some $i$, and that $\exists s,t$ such that $s<i<t-1$, $s>1$ or $t<2n+r$, and $\beta(s)=t$. Also suppose the leftmost anchor is at some index $u>t$. Then we have a sub-matching $\beta(s,t)$, and by lemma 3.6 $E_\beta\subset Y_\beta$. Then, using corollary 3.2, the following equality holds:

	$$(1+q)x_\beta(q)+\sum_{\alpha\in E_\beta,\alpha\not=\beta}q^{1/2}x_\alpha(q)=$$

	$$(x_{\beta(1,s-1)}(q)\frac{Q_{n+r-1}...Q_{n+r-(s-1)/2}}{Q_1...Q_{(s-1)/2-1}})((1+q)x_{\beta(s,2n+r)}(q)+\sum_{\alpha\in E_\beta,\alpha\not=\beta}q^{1/2}x_{\alpha(s,2n+r)}(q))=$$

	$$(x_{\beta(t+1,2n+r)}(q)\frac{Q_{n+r-(s-1)/2-1}...Q_{n+r-t/2}}{Q_1...Q_{(t-s+1)/2-1}})(x_{\beta(1,s-1)}(q)\frac{Q_{n+r-1}...Q_{n+r-(s-1)/2}}{Q_1...Q_{(s-1)/2-1}})((1+q)x_{\beta(s,t)}(q)+\sum_{\alpha\in E_\beta,\alpha\not=\beta}q^{1/2}x_{\alpha(s,t)}(q))$$

	We have that $e>j$ for any $Q_j$ term appearing in the equation above, and $e>n'+r'$ for any sub-matching coordinate appearing above, so by corollary 3.3:

	$$(1+q)x_\beta(q)+\sum_{\alpha\in E_\beta,\alpha\not=\beta}q^{1/2}x_\alpha(q)=0$$

	if and only if

	$$(1+q)x_{\beta(s,t)}(q)+\sum_{\alpha\in E_\beta,\alpha\not=\beta}q^{1/2}x_{\alpha(s,t)}(q)=0$$

	Note that $(\beta(s,t))(1)=t-s+1$. So by our inductive hypothesis (ii), we have

	$$(1+q)x_{\beta(s,t)}(q)+\sum_{\alpha\in E_{\beta(s,t)},\alpha\not=\beta(s,t)}q^{1/2}x_{\alpha}(q)=0$$

	By lemma 3.5, if $\alpha\in Y_\beta$, $(1+T_i)\alpha=q^{1/2}\beta$ if and only if $(1+T_{i-s+1})\alpha(s,t)=q^{1/2}\beta(s,t)$, so the previous equation implies $$(1+q)x_{\beta(s,t)}(q)+\sum_{\alpha\in E_\beta,\alpha\not=\beta}q^{1/2}x_{\alpha(s,t)}(q)=0$$

	as desired, and this case is proved.

	\vspace{5mm}
	\item Again take $\beta\in (1+T_i)W_{2n+r}^r$ for some $i$, but suppose the leftmost anchor is at some position $u$ where $1<u<i$. Then, as before, we have a sub-matching $\beta(u,2n+r)$ and by lemma 3.6 $E_\beta\subset Y_\beta$.

	Note that in both corollary 3.2 and our inductive hypothesis (ii) we specified cases involving anchors. This allows the exact same logic from the proof of the first case to prove this case.

	\vspace{5mm}
	\item Suppose $\beta\in (1+T_i)W_{2n+r}^r$ for some $i$, the leftmost anchor is at a position $u>i+1$, and $\not\exists s,t$ such that $\beta(s)=t$ and $s<i<t-1$. Lemma $3.7$ characterizes all $\alpha\in E_\beta$. We would like to prove the following for arbitrary $q$ where $e>n+r$:

	$$(1+q)x_\beta(q)+\sum_{\alpha\in E_\beta,\alpha\not=\beta}q^{1/2}x_\alpha(q)=-q^{1/2}x_{\beta(1,i-1)}(q)x_{\beta(i+2,2n+r)}(q)\frac{Q_{n+r}...Q_{n+r-(i-1)/2}}{Q_1...Q_{(i-1)/2}}$$

	See figure 9 for an example of this equality.
	\\

	We will prove this equality through yet another inductive proof, this time inducting on the number of top level humps to the left of the rightmost anchor.

	Formally, as we have in earlier lemmas, we will define $a_j$ by $a_1:=\beta(1)$, $a_j:=\beta(a_{j-1}+1)$. Then $a_b=u$ for some $b>1$.

	\end{enumerate}
\end{proof}
